---
title: "Final Report"
author: "Team 6 purple"
date: "9/17/2020"
output:
  pdf_document: default
  html_notebook: default
---

# PART II -- 

## Summary

## Introduction

It is common to believe that job training is efficient in boosting the earnings of disadvantaged workers. In 1970s, several experiments was conducted to explore the real impact of job training on wages, including the National Supported Work (NSW) Demonstration[1]. In the experiment, eligible workers were assigned to receive job training, and their incomes in 1973, 1974, and 1977 were recorded accordingly.

This analysis's primary goal is to assess whether receiving job training has a significant result in the possibility of receiving a non-zero wage. Other considerations include whether this effect differs by demographic groups in terms of age, education, and racial identity. We're also interested in exploring other associations with positive income.

**Data Processing**

The original data set is accessed via the NSW Demonstration experiment with 614 non-empty observations. Since this analysis features the possibility of getting a paid job, we create a binary variable 'inc78' where 0 represents zero wages, and 1 indicates positive income. The factorized variable 'inc78_factor' will be used as our response variable for analysis and modeling. The cleaned dataset included 10 distinctive predictors -- 5 categorical ones and 4 numeric ones.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(out.width = '50%')
knitr::opts_chunk$set(fig.align="center")
```

```{r library, message=FALSE}
library(tidyverse) # Data wrangling
library(ggplot2) # Plotting

# For logistic regression
library(arm)
library(car)
library(pROC)
library(e1071)
library(caret)

## table
library(knitr)
library(kableExtra)
```

```{r import data}
income = read.table("lalondedata.txt", header = T, sep = ",", dec = ".")
```

```{r data processing}
# Change the variable to factor levels
income = income %>% mutate(
        inc78 = case_when(
                re78 == 0 ~ 0,
                re78 > 0 ~ 1, )
        )

income$inc78_factor <- as.factor(income$inc78)
income$treat <- as.factor(income$treat)
income$black <- as.factor(income$black)
income$hispan <- as.factor(income$hispan)
income$married <- as.factor(income$married)
income$nodegree <- as.factor(income$nodegree)
```

## Exploratory Data Analysis

EDA is a process of uncovering useful information and pattern related to non-zero wages by statistical summary, table, and visualization.

**Response_Variable**

To get a general understanding of the baseline probability for model fitting and further analysis and examine the sufficiency of observations for both levels, we conduct a table of the response variables 'inc78_factor'. According to the table, among the 614 participants, the average possibility of getting a non-zero wage job in 1978 is around 76%.

```{r response varible}
table(income$inc78_factor)
```

**Treatment vs Non-zero Wage**

As an independent predictor, the treatment could be a reasonable predictor for predicting the possibility of having a non-zero income. Given the condition where a worker received job training, the conditional probability of getting positive pay decreases slightly compared to those who did not. However, according to the result of chi-square test (p-value = 0.77), the difference is not significant. Therefore, we need to exam the effect of treatment and possible interaction effects in the final model.

```{r}
apply(table(income[,c("inc78_factor","treat")])/sum(table(income[,c("inc78_factor","treat")])), 2, function(x) x/sum(x))
chisq.test(table(income[,c("inc78_factor","treat")]))
```

**Other Categorical Variables**

Regarding other categorical variables, the association between Black and Non-zero income seems to be significant according to the conditional probability table. There's a lower conditional possibility of having non-zero wage when the worker's racial identity is Black, compared to workers in other racial groups. Even though there's a difference between the non-Hispanic and Hispanic groups, the change could result from insufficient observations on Hispanic workers. Since no significant change in the possibilities cross different levels is observed in 'married' and 'nondegree,' they might be pool predictor for predicting the non-zero wage rate. Still, possible interaction effects will be tested later. 

```{r}
apply(table(income[,c("inc78_factor","black")])/sum(table(income[,c("inc78_factor","black")])),
      2,function(x) x/sum(x))

apply(table(income[,c("inc78_factor","hispan")])/sum(table(income[,c("inc78_factor","hispan")])),
      2,function(x) x/sum(x))
```

**Numeric Variables**

According to the boxplot analysis, the mean and range of age in the positive earning group is significantly higher zero income group. Given a chi-square test result of 0.002, we suggest a reasonable association between age and a positive wage. The possibility of getting non-zero earning decreases as the worker gets older, keeping everything else the same. However, considering the small sample size in different age groups, the effect of age on non-zero will be re-examed latter in modeling selection.

```{r}
ggplot(income,aes(x = inc78_factor, y = age)) +
  geom_boxplot()
```

**Interaction Effect**

Interaction effects occur when the impact of one variable depends on the value of another variable. Conditional possibility tables and faceted boxplots are used to test the interaction effect between terms. 

Given whether the participants received job training, the visualized change on the correlation between non-zero wage and age indicates a possible interaction effect between treatment and age. The benefit of job training increases as the worker increases in age, and the net effects of training and age on the possibility of getting a non-zero wage shift from negative to zero, and possibly positive after a certain age.

Another notable interaction is between age and degree as the association between inc78_factor and age change significantly, moving from groups with high school degrees to groups without. The possibility of getting paid employment tends to stay the same, and even increases as a worker with a high school degree grow older, while the possibility decrease as one without a high school degree.

To draw a solid conclusion on interaction terms' impact, we will determine their efficiency in the modeling selection process.

```{r}
ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~treat)

ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~nodegree)
```

## Model Selection 

The model selection is based on two methods: AIC Stepwise Selection and ANOVA chi-square test. 

Applying stepwise selection on the full model with all potential predictor and a null model containing only treatment as a predictor, we are able to combine the forward and backward selection and pick the most significant variables by either adding or removing variables several times to meet the maximum likelihood estimation. At the end of the process, two significant variables -- age and black -- are selected. Stepwise selections manage large amounts of potential predictor variables and provide us a concise regression that allows us to add individual predictors and interaction effects later through the ANOVA test. 

Interaction terms between treatment and other variables were added to the full model one by one to explore the effect of demographic groups on the association between non-zero wage and job training. Given a p-value of 0.038 when we use ANOVA to compare the model with interactions between age and treatment vs. without, we reject the null hypothesis and confirm the interaction term efficient for predicting the possibility of getting a non-zero wage. 

Even though the interaction effect between age and high school degree seems significant in previous EDA, we failed to reject the null hypothesis and found the interaction term negligible. Other individual predictors such as nodegree, married and Hispanic have poor performance on ANOVA test, indicating insignificant effect on non-zero wages.



**Final Model**

#Add the final model HERE#

As the model summary statistics suggested, Black, treatment, age and the interaction terms between age and treatment are four significant variables in the final model, using a threshold of 0.1. The residual deviance of the final model is 666.50 which is significantly higher from the null deviance 643.51, thus suggesting model sufficient in predicting the possibility of getting non-zero wage.

*Association suggested by the final model*

  -- Keeping other variables the constant, compared to people in the treatment group, the average odds possibility of getting paid jobs for those without job training decrease by 72% with a 90% range of [7%, 92%]. However, the effect of treatment on the non-zero wage rate does changes according to ages as a worker get 2.3% net increase in the odds of getting non-zero wage, compared to people without training. Therefore, the benefit of job training increases as the worker increases in age, and the net effects of training and age on the possibility of getting a non-zero wage will be positive after a certain age.
 
 -- There's a strong association between age and the possibility of positive wage. Keep other variables the constant, one unit increase in age will result in a 3.9% decrease in the average odds possibility of getting non-zero wage.
 
 -- Keeping everything else the sames, the odds of getting non-zero wages will drop by 52.9% if the worker's racial identity is Black, compared to worker in other racial groups.

```{r}
final <- glm(inc78_factor~black+treat*age,family = binomial,data = income)
summary(final)
exp(final$coefficients)
exp(confint(final, 'treat1', level=0.90))
```

## Model Accessment

For the model assessment of existed model, we calculate the raw (response) residuals for fitted logistic regression and plot average residual versus average predicted probability (or average predictor value) for each bin to explore abnormal patterns. According to the binned residual plot, all the points fall between the red lines which represent a band expected 95% of the observations. Therefore, we confirmed the model assumption is not violated. 

Moreover, cook's distance is used to examine any potential outliers in the model. [........]

Finally, variance inflation factors was used to confirm that no multicolinearity existed in the chosen model. As all the VIFs were confirmed to be below 5 except the interaction terms, the model was finalized and used for inference.

```{r binned residual, echo = FALSE}
residFinal <- residuals(final,"resp")
binnedplot(x=fitted(final),y=residFinal,xlab="Pred. probabilities",col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy") # 92% within SE band
```

```{r collinearity}
vif(final)
```

## Model Validation

For model validation, we used a 2x2 confusion matrix for evaluate the performance of out final model. According to the matrix, we got an accuracy of 0.77 which means 77% of the observation are correctly classified by the model. The sensitivity of the test reflects the probability that the prediction will be positive (having non-zero wage) among those who do have non-zero earnings. A 0.99 sensitivity indicates a great true-positive rate of the model. However, the model is less efficient for giving a correct result when the worker does not have a positive earnings since the specificity rate is very low.

To better understand the performance of the classifier over all possible thresholds, we generated a ROC curve by plotting the True Positive Rate (y-axis) against the False Positive Rate (x-axis). The highest accuracy we can observed from adjusting the threshold for assigning observations to the given class is 0.757.

```{r}
# Consider deleting this chunk
# Threshold = 0.5
Conf_mat <- confusionMatrix(as.factor(ifelse(fitted(final) >= 0.5, "1","0")),
                            as.factor(income$inc78_factor),positive = "1")
Conf_mat$table
Conf_mat$overall["Accuracy"];
Conf_mat$byClass[c("Sensitivity","Specificity")] #True positive rate and True negative rate
```

**ROC curve: final model vs model with all predictors**

```{r warning=FALSE,message=FALSE}
# ROC curve
invisible(roc(income$inc78_factor,fitted(final),plot=T))
invisible(roc(income$inc78_factor,fitted(raw),plot=T))
```

## Conclusion and Limitation

[..........]

## Appendix and Reference
[1] Lalonde, R. J. (1986), Evaluating the econometric evaluations of training programs with experimental data, The American Economic Review, 76, 604 - 620