---
title: "Team proj 1"
author: "Team 6 purple"
date: "9/17/2020"
output: html_document
---

### ---Environment setup, delete this caption for final report---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(out.width = '50%')
knitr::opts_chunk$set(fig.align="center")
```

```{r library,message=FALSE}
library(tidyverse) # Data wrangling
library(ggplot2) # Plotting

# For logistic regression
library(arm)
library(car)
library(pROC)
library(e1071)
library(caret)
```

### ---Data wrangling
```{r import data}
# CHANGE DIRECTORY
income <- read.table("C:/Users/renha/Desktop/IDS702/lalondedata.txt",header = TRUE,sep = ",",stringsAsFactors = FALSE,row.names = 'X')
#income = read.table(file = "D:/MIDS/fall 2020/702 modeling/dataset/lalondedata.txt", header = T, sep = ",", dec = ".")
```

```{r}
# Change the variable to factor levels
income = income %>% mutate(
        inc78 = case_when(
                re78 == 0 ~ 0,
                re78 > 0 ~ 1 )
        )

# Change to factors, apply to part I and part II
income$treat <- as.factor(income$treat)
income$black <- as.factor(income$black)
income$hispan <- as.factor(income$hispan)
income$married <- as.factor(income$married)
income$nodegree <- as.factor(income$nodegree)

# Response variable for part I, change in income - continuous
income$change <- income$re78 - income$re74

# Response variable for part II, inc78_factor - binary
income$inc78_factor <- as.factor(income$inc78)
```

## Part I

### EDA

#### Response variable - change in income

```{r}
hist(income$change)
summary(income$change)
```

#### Response variable vs continous variables
```{r}
# continuous
pairs(income[,-c(1,4,5,6,7,11)])# Multicollinearity check?
# Could add a line y=0
plot(income$age,income$edu)# ?? examine interaction?
plot(income$age,income$change)
plot(income$educ,income$change)
```

#### Response variable vs categorical variables, interactions:

```{r}
ggplot(data= income, aes(x=treat, y=change)) + geom_boxplot() 

ggplot(data= income, aes(x=as.factor(educ), y=change)) + geom_boxplot() + facet_wrap(~treat)
ggplot(data= income, aes(x=black, y=change)) + geom_boxplot() + facet_wrap(~treat) # interaction black:treat
ggplot(data= income, aes(x=hispan, y=change)) + geom_boxplot() + facet_wrap(~treat) 
ggplot(data= income, aes(x=married, y=change)) + geom_boxplot() + facet_wrap(~treat) # interaction married:treat
# ggplot(data= income, aes(x=nodegree, y=change)) + geom_boxplot() + facet_wrap(~treat) 
ggplot(data= income, aes(x=treat, y=change)) + geom_boxplot() + facet_wrap(~nodegree) # nodegree:treat 
ggplot(data= income, aes(x=age, y=change)) + geom_point() + geom_smooth(method = 'lm') + facet_wrap(~treat) # age:treat
```

### Model 

```{r}
raw <- lm(change~treat+age+educ+black+hispan+married+nodegree,data = income)
null <- lm(change~treat,data=income)
step(null,scope=formula(raw),direction="both",trace=0)
```

```{r}
fit1 <- lm(change~treat+age+married, data = income)
fit2 <- lm(change~treat+age+married+nodegree+nodegree:treat, data = income)
anova(fit1,fit2)
fit3 <- lm(change~treat+age+married+married:treat, data = income)
anova(fit1,fit3)
fit4 <- lm(change~treat+age+married+black, data = income)
anova(fit1,fit4)
fit5 <- lm(change~treat+age+married+black+black:treat, data = income)
anova(fit1,fit5)
fit6 <- lm(change~treat+age+married+nodegree, data = income)
anova(fit1,fit6)
fit7 <- lm(change~treat+age+married+treat:age, data = income)
anova(fit1,fit7)
```

```{r}
vif(fit7)
ggplot(data = income,aes(x=age,y = fit7$residuals))+
         geom_point()+
         geom_smooth(method = 'lm')
plot(fit7, which = 1)
plot(fit7, which = 2)
n <- nrow(model.matrix(fit7))
p <- ncol(model.matrix(fit7))
lev_scores <- hatvalues(fit7)
plot(lev_scores,col=ifelse(lev_scores > (2*p/n), 'red2', 'navy'),type="h",
     ylab="Leverage score",xlab="Index",main="Leverage Scores for all observations")
text(x=c(1:n)[lev_scores > (2*p/n)]+c(rep(2,4),-2,2),y=lev_scores[lev_scores > (2*p/n)],
     labels=c(1:n)[lev_scores > (2*p/n)])
plot(fit7,which=4,col=c("blue4"))
threshhold = 2*(p+1)/n
abline(h=threshhold)
#plot(fit7,which=5,col=c("blue4"))
```

```{r}
summary(fit7)
confint(fit7, level = 0.9)
#confint(fit7, level = 0.95)
```


```{r}
#cleaned <- income[-c(132,79,334),]
#new <- lm(change~treat+age+married+treat:age, data = cleaned)
#summary(new)
#vif(new)
```

```{r}
set.seed(702) 
income <- income[sample(nrow(income)),]
K <- 10
RMSE1 <- matrix(0,nrow=K,ncol=1)
RMSE2 <- matrix(0,nrow=K,ncol=1)
kth_fold <- cut(seq(1,nrow(income)),breaks=K,labels=FALSE)
for(k in 1:K){
        test_index <- which(kth_fold==k)
        train <- income[-test_index,]
        test <- income[test_index,]
        test.fit1 <- lm(change~treat+age+married+treat:age, data = train)
        test.fit2 <- lm(change~treat+age+married, data = train)
        y_test_pred1 <- predict(test.fit1,test)
        y_test_pred2 <- predict(test.fit2,test)
        RMSE1[k,] <- sqrt(mean((test$change - (y_test_pred1))^2))
        RMSE2[k,] <- sqrt(mean((test$change - (y_test_pred2))^2))
}
mean(RMSE1)
mean(RMSE2)
```

## Part II

### EDA

#### Response variable vs categorical variables

```{r}
apply(table(income[,c("inc78_factor","treat")])/sum(table(income[,c("inc78_factor","treat")])),
      2,function(x) x/sum(x))
apply(table(income[,c("inc78_factor","black")])/sum(table(income[,c("inc78_factor","black")])),
      2,function(x) x/sum(x))
apply(table(income[,c("inc78_factor","hispan")])/sum(table(income[,c("inc78_factor","hispan")])),
      2,function(x) x/sum(x))
apply(table(income[,c("inc78_factor","married")])/sum(table(income[,c("inc78_factor","married")])),
      2,function(x) x/sum(x))
apply(table(income[,c("inc78_factor","nodegree")])/sum(table(income[,c("inc78_factor","nodegree")])),
      2,function(x) x/sum(x))
apply(table(income[,c("inc78_factor","educ")])/sum(table(income[,c("inc78_factor","educ")])),
      2,function(x) x/sum(x))
```

#### Response variable vs discrete variables, interaction with other variables

```{r, echo=FALSE}
# Check interactions between age and other predictors
ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~treat)
ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~black)
ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~hispan)
ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~married)
ggplot(data= income, aes(x=age, y=inc78_factor)) + geom_boxplot() + facet_wrap(~nodegree) # possible nodegree:age
```


#### Chi-square test for independence
```{r}
chisq.test(table(income[,c("nodegree","age")]))
chisq.test(table(income[,c("treat","educ")]))
chisq.test(table(income[,c("black","educ")]))
chisq.test(table(income[,c("married","educ")]))
chisq.test(table(income[,c("hispan","inc78_factor")])) # independent hispan inc78_factor
chisq.test(table(income[,c("black","inc78_factor")]))
chisq.test(table(income[,c("married","inc78_factor")])) # independent
chisq.test(table(income[,c("nodegree","inc78_factor")])) # independent

chisq.test(table(income[,c("hispan","age")])) # independent hispan age
chisq.test(table(income[,c("hispan","educ")]))
chisq.test(table(income[,c("hispan","black")]))
chisq.test(table(income[,c("hispan","nodegree")]))
chisq.test(table(income[,c("hispan","married")])) # independent hispan married
```

### Model

#### Model Selection: AIC
```{r}
full <- glm(inc78_factor~treat+age+educ+black+hispan+married+nodegree+nodegree:age+treat:black+treat:hispan,family = binomial,data = income)
summary(full)
residFull <- residuals(full,"resp")
binnedplot(x=fitted(full),y=residFull,xlab="Pred. probabilities",col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy") # 92% within SE band
```


```{r}
raw <- glm(inc78_factor~treat+age+educ+black+hispan+married+nodegree,family = binomial,data = income)
summary(raw)
```


```{r}
resid <- residuals(raw,"resp")
binnedplot(x=fitted(raw),y=resid,xlab="Pred. probabilities",col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy") # 92% within SE band
```

```{r}
null <- glm(inc78_factor~treat,data=income,family=binomial)
selection <- step(null,scope=formula(full),direction="both",trace=0)
```

#### Model Selection: ANOVA Chi-square test
```{r}
fit1 <- glm(inc78_factor~treat+age+black,family = binomial,data = income)
anova(selection,fit1,test = 'Chisq')

fit2 <- glm(inc78_factor~age+treat*black,family = binomial,data = income)
anova(fit2,selection,test = 'Chisq')

fit3 <- glm(inc78_factor~age+black+treat*hispan,family = binomial,data = income)
anova(fit3,selection,test = 'Chisq')

fit4 <- glm(inc78_factor~black+treat*age,family = binomial,data = income)
anova(fit4,selection,test = 'Chisq') # treat:age significant

fit5 <- glm(inc78_factor~age+black+treat*educ,family = binomial,data = income)
anova(fit5,selection,test = 'Chisq')

fit6 <- glm(inc78_factor~age+black+treat*married,family = binomial,data = income)
anova(fit6,selection,test = 'Chisq')

fit7 <- glm(inc78_factor~age+black+treat*nodegree,family = binomial,data = income)
anova(fit7,selection,test = 'Chisq')

fit8 <- glm(inc78_factor~treat+black+age*nodegree,family = binomial,data = income)
anova(fit8,fit7,test = 'Chisq')

fit9 <- glm(inc78_factor~black+age*nodegree,family = binomial,data = income)
anova(fit9,selection,test = 'Chisq')
```

#### Final Model
```{r}
final <- glm(inc78_factor~black+treat*age,family = binomial,data = income)
#final_ploy <- glm(inc78_factor~black+treat*age+poly(age,3),family = binomial,data = income)
vif(final)
summary(final)
exp(final$coefficients)
exp(confint(final,level = 0.9))
```

#### Model assessment: binned residual plot
```{r}
residold <- residuals(final,"resp")
binnedplot(x=fitted(final),y=residold,xlab="Pred. probabilities",col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy") # 92% within SE band -> 95.8%
binnedplot(x=income$age,y=residold,xlab="Age",col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")
```

##### Add polynomial term - for reference
```{r include=FALSE}
#residnew <- residuals(final_poly,"resp")
#binnedplot(x=fitted(final_poly),y=residold,xlab="Pred. probabilities",col.int="red4",ylab="Avg. #residuals",main="Binned residual plot",col.pts="navy") # 92% within SE band -> 95.8%
#binnedplot(x=income$age,y=residnew,xlab="Age",col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")
```

#### Model validation: Confusion matrix, Sensitivity, Specificity, and ROC curve

##### Threshold = 0.5, final model
```{r}
# Threshold = 0.5
anova(raw,final,test = 'Chisq')
Conf_mat <- confusionMatrix(as.factor(ifelse(fitted(final) >= 0.5, "1","0")),
                            as.factor(income$inc78_factor),positive = "1")
Conf_mat$table
Conf_mat$overall["Accuracy"];
Conf_mat$byClass[c("Sensitivity","Specificity")] #True positive rate and True negative rate
```

##### Threshold = 0.5, model with polynomial term, model with all predictors
```{r include=FALSE}
# Threshold = 0.5
#anova(final,final_poly,test = 'Chisq')
#Conf_mat <- confusionMatrix(as.factor(ifelse(fitted(newfinal) >= 0.5, "1","0")),
#                            as.factor(income$inc78_factor),positive = "1")
#Conf_mat$table
#Conf_mat$overall["Accuracy"];
#Conf_mat$byClass[c("Sensitivity","Specificity")] #True positive rate and True negative rate

#Conf_mat <- confusionMatrix(as.factor(ifelse(fitted(raw) >= 0.5, "1","0")),
#                            as.factor(income$inc78_factor),positive = "1")
#Conf_mat$table
#Conf_mat$overall["Accuracy"];
#Conf_mat$byClass[c("Sensitivity","Specificity")] #True positive rate and True negative rate
```

##### Threshold = mean-level, final model
```{r, include=F}
# Consider deleting this chunk
# Threshold = mean level
Conf_mat <- confusionMatrix(as.factor(ifelse(fitted(final) >= mean(income$inc78), "1","0")),
                            as.factor(income$inc78_factor),positive = "1")
Conf_mat$table
Conf_mat$overall["Accuracy"];
Conf_mat$byClass[c("Sensitivity","Specificity")] #True positive rate and True negative rate
```

##### ROC curve: final model vs model with all predictors
```{r warning=FALSE,message=FALSE}
# ROC curve
invisible(roc(income$inc78_factor,fitted(final),plot=T,print.thres="best",print.auc=T,legacy.axes=T,col="red3"))
invisible(roc(income$inc78_factor,fitted(raw),plot=T,legacy.axes=T,add=T,col="blue3"))
legend('bottomright', c('final','raw'),lty=c(1,1),
       lwd=c(2,2),col=c('red3','blue3'))
```

